\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-15 paper.tex
save

We assume that we already have an accurate simulation of the stochastic
generative process that defines $p_\mathbf{\theta}(\mathbf{x}|\mathbf{z})$, as
specified through a deterministic function $g(\cdot; \theta) : \mathbb{R}^m \to
\mathbb{R}^d$. That is
\begin{equation}\label{eqn:p_x_sim}
    p_\mathbf{\theta}(\mathbf{x}) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\mathbf{z}:g(\mathbf{z};\mathbf{\theta}) \leq \mathbf{x}\}} p(\mathbf{z}) d\mathbf{z}.
\end{equation}
The simulator $g$ is assumed to be a non-invertible function, that can only be
used to generate data in forward mode. For this reason, evaluating the integral
in Eqn.~\ref{eqn:p_x_sim} is intractable. Importantly, and as increasingly found
in science, we consider the additional constraint that $g$ is a
non-differentiable model, e.g. when specified as a computer program.

@@

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound $U(\psi)$ is differentiable, and using the log-likelihood trick it comes:
\begin{align}\label{eqn:vo-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \int f(\mathbf{c})  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int f(\mathbf{c}) \nabla_\psi q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int \left[ f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c}) \right]  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c} \nabla_\psi \log q_\psi(\mathbf{c}))]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log q_\psi(\mathbf{c})$ of the proposal
is known and that one can evaluate $f(\mathbf{\mathbf{c}})$ for any $\mathbf{c}$, then
one can construct empirical estimates of Eqn.~\ref{eqn:vo-grad}, which can
in turn be used to perform stochastic gradient descent (or a variant thereof)
in order to minimize $U(\psi)$.

@@

>>>>

We assume that we already have an accurate simulation of the stochastic
generative process that defines $p_\mathbf{\theta}(\mathbf{x}|\mathbf{z})$, as
specified through a deterministic function $g(\cdot; \theta) : \mathbb{R}^m \to
\mathbb{R}^d$. That is
\begin{equation}\label{eqn:p_x_sim}
    p_\mathbf{\theta}(\mathbf{x}) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\mathbf{z}:g(\mathbf{z};\mathbf{\theta}) \leq \mathbf{x}\}} p(\mathbf{z}) d\mathbf{z}.
\end{equation}
The simulator $g$ is assumed to be a non-invertible function, that can only be
used to generate data in forward mode. For this reason, evaluating the integral
in Eqn.~\ref{eqn:p_x_sim} is intractable. Importantly, and as commonly found
in science, we finally assume the lack of access to or existence of derivatives of $g$ with respect to $\theta$, e.g. when specified as a computer program.

@@

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound $U(\psi)$ is differentiable, and using the log-likelihood trick it comes:
\begin{align}\label{eqn:vo-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c})] \nonumber \\
    &= \nabla_\psi \int f(\mathbf{c})  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int f(\mathbf{c}) \nabla_\psi q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int \left[ f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c}) \right]  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c})]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log q_\psi(\mathbf{c})$ of the proposal
is known and that one can evaluate $f(\mathbf{\mathbf{c}})$ for any $\mathbf{c}$, then
one can construct empirical estimates of Eqn.~\ref{eqn:vo-grad}, which can
in turn be used to perform stochastic gradient descent (or a variant thereof)
in order to minimize $U(\psi)$.

@@

\glnote{Naive approach by using VO on ${\cal L}_g$.}
\glnote{Smarter approach by exploiting the fact that $\partial d / \partial x$ is known exactly.}
